Verification of the CVM algorithm with a New Recursive Analysis Technique

### 2025-02-18 paper/sec/invariants.tex
slight edit

As a simpler concrete example, consider choosing $S = \{s\}$ for any element $s \in A$.
If we pick $h_1(x) = x$ and $h_2(x) = M+1-x$ (for sufficiently large $M$), we obtain from~\cref{i:func_invariant}:
\begin{align*}
\expect \left[ p^{-1} X_s \right] &= \expect \left[ \prod_{s \in S} h_1(p^{-1} X_s) \right] \leq h_1(1)^{|S|} = 1\\
\expect [M + 1 - p^{-1} X_s] &= \expect \left[ \prod_{s \in S} h_2(p^{-1} X_s) \right] \leq h_2(1)^{|S|} = M
\end{align*}
This implies the simple invariant $\expect \left[ p^{-1} X_s \right] = 1$ (and thus, unbiasedness) without needing a separate induction over the algorithm.
The main takeaway here is that it is possible to derive useful and general probabilistic invariants by considering expectations of classes of functions of the state, proved using recursion or induction over the algorithm itself.
As far as we know this method of establishing tail bounds for randomized algorithms is new.

>>>>

As a simpler concrete example, consider choosing $S = \{s\}$ for any element $s \in A$.
If we pick $h_1(x) = x$ and $h_2(x) = M+1-x$ (for sufficiently large $M$), we obtain from~\cref{i:func_invariant}:
\begin{align*}
\expect \left[ p^{-1} X_s \right] &= \expect \left[ \prod_{s \in S} h_1(p^{-1} X_s) \right] \leq h_1(1)^{|S|} = 1\\
\expect [M + 1 - p^{-1} X_s] &= \expect \left[ \prod_{s \in S} h_2(p^{-1} X_s) \right] \leq h_2(1)^{|S|} = M
\end{align*}
This implies the simple invariant $\expect \left[ p^{-1} X_s \right] = 1$ without needing a separate induction over the algorithm.
The main takeaway here is that it is possible to derive useful and general probabilistic invariants by considering expectations of classes of functions of the state, proved using recursion or induction over the algorithm itself.
As far as we know this method of establishing tail bounds for randomized algorithms is new.
